---
title: "Identifying Schools with Biggest Differences in Math Test Scores, Part 2"
author: "Aicha Bangoura"
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
options(scipen = 999)
```

### Introduction

We'll do what we did in class, but with a twist: instead of Montgomery County, Pa., you'll be working with data from Montgomery County, Md. Same basic idea, though: linear regression looking at free and reduced-price meals and math test scores.

Just as in the exercise, you will create a scatterplot to visualize the relationship between the FARMS (Free and Reduced-price Meal System) percentage and the percentage of students who met or exceeded math expectations. You'll then use linear regression to identify the schools that show the biggest differences between their actual and expected math test scores and evaluate the results of your model.

### Loading the Data

First, let's load the data into a data frame using the read_csv() function from the tidyverse library.

```{r}
schools <- read_csv("https://raw.githubusercontent.com/dwillis/jour405_files/refs/heads/main/montgomery_md.csv")
```
### Task: Create a Scatterplot and Describe the Results (5 points)

First, let's create a scatterplot to visualize the relationship between the FARMS percentage and the percentage of students who met or exceeded math expectations. We'll use ggplot2 to create the plot and add a linear regression line.

```{r}
schools |>
  ggplot(aes(x = `FARMS Pct`, y = `% Met or Exceeded Math Expectations`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "FARMS Percentage", y = "% Met or Exceeded Math Expectations", title = "Scatterplot of Math Expectations vs FARMS Percentage")
```
Based on what you see here, what's the relationship between FARMS percentage and math test scores? Compare this scatterplot to the one you made in the class activity.

ANSWER HERE: This scatter plot shows a  strong-negative relationship between FARMS and math scores. The scatter plot we were shown in class was a weak-positive relationship.


### Task: Linear Regression Model (7 points)

Next, you'll perform a linear regression analysis using the lm() function. We'll set the percentage of students who met or exceeded math expectations as the dependent variable, and the FARMS percentage as the independent variable. In other words, if we know the FARMS percentage for a school, how reliably can we predict its math score?

```{r}
model <- lm(`% Met or Exceeded Math Expectations` ~ `FARMS Pct`, data = schools)

summary(model)
```

Provide the model's output for the three statistics mentioned in the class activity.

ANSWER HERE: The FARMS percentage is -0.595 which is the true value. The multiple R- squared is 0.676. Residual standard error is 11.04 meaning the predictions were off by about 11%. 

Based on what you've learned so far, what does it tell us about the nature of the relationship between FARMS percentage and math scores? Is this model worth exploring further? Why or why not?

ANSWER HERE: From what I've learned so far, this tells me that the relationship between FARMS percentage and math scores is negative since the slope goes down. Yes, this model is worth exploring further since there's enough evidence which links poverty and low scores. More in-depth investigation is required to confirm the correlation between the two. 

### Task: Identifying Schools with Biggest Residuals (3 points)

Let's calculate and examine the residuals next. 

```{r}
schools$predicted <- predict(model)

# Calculate the residuals
schools <- schools |> 
  mutate(residual = `% Met or Exceeded Math Expectations` - predicted)

# Sort the data frame by the absolute value of the residuals in descending order
schools <- schools |>
  mutate(abs_residual = abs(residual)) |>
  arrange(desc(abs_residual)) 

```

Look at the absolute values of the residuals; what are the largest differences between the predicted result and the actual math scores? Pick two schools with the largest residuals. Find the zip codes where they are located and using the CensusReporter.org website, describe the most interesting demographic data from each zip code.

ANSWER HERE: The largest differences between the predicted results and the actual math scores are as follows: WHITEMARSH Elementary School with a predicted math score of 71.93% and an absolute score of -45.83%. The second largest difference is LOWER GWYNEDD Elementary School with a predicted math score of 63.40% and an  absolute score of -31.99%. The two schools with the largest residuals are CROOKED BILLET EL SCH with 30.18% and WOODLAND EL SCH with 25.38%.


## Task: Evaluate the model (5 points)

Describe what conclusions you'd draw from this model's results and how it would influence a story pitch about the school board member and math scores.

ANSWER HERE: The conclusion I could draw on are that schools were students display need is where performance in tests are low. The models suggests that while the low-income population in schools perform poorly of math scores, it doesn't show a direct relationship between Free and reduced meals for students and their scores. A story I could pitch is that the school board should implement more than free and reduced meals to everyone in the school, such as counseling for mental and emotional health and other assistance.
