---
title: "HW4 MD CityCrime"
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse)
#install.packages ("janitor")
library(janitor)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
crimes <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv")
```
### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}
md_cities_rates <- crimes |>
  mutate(violent_rate_2019 = (violent2019 / pop2019) * 100,
    property_rate_2019 = (property2019 / pop2019) * 100,
    violent_rate_2020 = (violent2020 / pop2020) * 100,
    property_rate_2020 = (property2020 / pop2020) * 100)

```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?

ANSWER HERE: Baltimore is the top for the crimes reported as in 2019 and 2020 the violent and property crimes outnumber any other city or county. It falls in the highest populated area so that seems to influence it. The smallest city in Maryland is Eagle Harbor with 65 being listed as the population. It is also the cleanest city for no crimes reported. The story is that the higher the population, the higher the crimes rates and vice versa. 


### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
 md_cities_change <- crimes |> mutate(
   violent_rate_2020 = (violent2020 - violent2019) / violent2019 * 100,
   property_rate_2019 = (property2020 - property2019) / property2019 * 100
   )

```

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?

ANSWER HERE: Baltimore's violent crime rate decreased by 15.4 percent for 2020 in which the population decreased in that year. The property rate also plunged by 28 percent. Prince George's county doesn't change in any way from the data. The data displays  incorporated cities including Rockville and Gaithersburg, which both increased in violent crimes. That information was interesting as I always see Montgomery County as a progressive county. 

### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the _best_ story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

ANSWER HERE: Maryland sees a crime rate decrease for several  cities in the year 2020 including Baltimore. Maryland's reputation for crime has been compromised for only a few cities including Ocean City, Salisbury, Laurel, Hyattsville and the notorious Baltimore. Shockingly enough the rates went down in the year that people arguably were the most angst. The worldwide pandemic saw people losing jobs and loved ones getting sick and in the most vulnerable times the violence and property crimes went down particularly in on of the most dangerous cities being Baltimore. This shows a small progress for people in the city as well as investment opportunites. 

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.

### Task 10: 
Using the 2020 violent crime rate, calculate the percentiles using `ntile()`. Save the results to a new dataframe called `md_cities_percentiles`. How many cities are above the 90th percentile?
```{r}
md_cities_percentiles <- md_cities_change |>
  mutate(violent_rate_percentile = ntile(violent_rate_2020, 100))
```
Answer: Ten cities are above the 90th percentile. 
```{r}
 md_cities_percentiles |> 
  summarize(iqr_violent_rate_2020 = IQR(violent_rate_2020, na.rm = TRUE))
```
Do the results change your best story idea from the previous assignment? Why or why not?
Answer Here:
The results don't change my "best story idea" from the previous assignment because Baltimore, Hyattsville, Salisbury, Ocean City and Laurel are all still in the lower levels of the violent rate percentile. The results make sense since more many of the cities rose in crime rate levels more versus the other cities and some stayed the same.



```


