---
---
title: "HW7_MandMs"
---

In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse
```{r}
library(tidyverse)
```


### Task 2: Load the M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv
```{r}
MandMs <- read_csv ("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")
```


### Task 3: Sample 5 rows of data from the dataset, and then sample 10 rows, saving each to separate dataframes called sample5 and sample10.
```{r}
sample_5 <- MandMs|>
sample_n(100) 
sample_10 <- MandMs |> 
sample_n(100)
```

### Task 4: Calculate the mean and standard deviation of the samples, using one of the colors.
```{r}
sample_5 |> summarize (mean = mean(red), sd =sd(red)) 

```



### Task 5: Calculate the mean and standard deviation of the population, using the same color as Task 4
```{r}
sample_10|> summarize (mean = mean(red), sd =sd(red)) 
```


### Task 6: Compare the sample and population statistics

Compare the three sets of means and standard deviations. Are they similar?
Answer Here: Yes, they're each similar with the mean being 72.98 and the standard deviation being 16.45

### Task 7: Describe the results

Write a few sentences about what you found in the data. What does this tell you about the M&Ms dataset and about the act of random sampling?
Answer Here: The data tells me that the average of red M&M's are around 73 and in terms of the standard deviation there are red M&M's that are above or lower than 73. The sample of 5 and 10 rows when there's 100 makes it less accurate. 


When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
