Aicha Bangoura

Some data I currently obtain consist of grades and gpa among students taking JOUR320 New Writing and Reporting with the several professors that teach it.

As soon as I complete my datasheet, I will run an ANOVA test for the final project. I could take on a One-way ANOVA Analysis with the data I have and compare the Grades and GPA distributed to students between professors. I could use the standard deviation formula. I could then run a post-anova Tukey test at the end to test the confidence intervals. If the p-value is higher or lower than 0.05, then my hypothesis will confirmed. 
```{r}
library(tidyverse)
```


```{r}
library(readr)

grades<-read_csv("Grades.csv")
```
```{r}
Summary(JOUR_320_grade_data_Sheet2_Sheet1)
```

# Original data

I'm able to load the data from the csv Google Sheet I made and view it on Environment, but I haven't got to a place to successfully run it yet. I just need to refresh from my previous works. Please let me know if the data needs further details.

```{r}
Summary("JOUR_320_grade_data_Sheet2_Sheet1")
```

```{r}
JOUR_320_grade_data_Sheet2_Sheet1 <- JOUR_320_grade_data_Sheet2_Sheet1 |> mutate()

```

